Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2015-08-31 paper.tex
Whitespace stuff
Trying out new title
Add m-dash and abstract

\title{Modeling vertical-axis cross-flow turbine wakes}

@@

\begin{abstract}
The wake of a vertical-axis cross-flow turbine (CFT) is modeled numerically via
several techniques.
\end{abstract}

@@

The flow field was modeled using the Reynolds-averaged Navier--Stokes equations,
employing two different turbulence models--Menter's $k$--$\omega$ SST
\cite{Menter1994} and the Spalart--Allmaras (SA) one equation model
\cite{Spalart1992}. The SST model was chosen due to its prominence in the
literature for simulating separating flows, which we assumed to be present in
the current problem in the form of dynamic stall. The SA model was shown by
Ferreira et al. \cite{Ferreira2007} to match experimental particle image
velocimetry (PIV) results for a CFT in dynamic stall, though this was a somewhat
low Reynolds number case ($5 \times 10^4$).

>>>>

\title{Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and
    3-D RANS}

@@

\begin{abstract}
The wake of a vertical-axis cross-flow turbine (CFT) is modeled numerically via
blade-resolved $k$--$\omega$ SST and Spalart--Allmaras RANS models in two and
three dimensions. Results for each case are compared with experimental
measurements of the turbine shaft power, overall drag, mean velocity, turbulence
kinetic energy, and momentum transport terms in the near-wake at one diameter
downstream. It is shown that 2-D simulations overpredict turbine loading and do
not resolve mean vertical momentum transport, which plays an important role in
the near-wake's momentum balance. The 3-D simulations fair better at predicting
performance, with the Spalart--Allmaras model predictions being closest to the
experiments. The SST model more accurately predicts the turbulence kinetic
energy while the Spalart--Allmaras model more closely matches the momentum
transport terms in the near-wake. These results show the potential of RANS as a
design tool and a way to ``interpolate'' experimental flow field measurements.
\end{abstract}

@@

The flow field was modeled using the Reynolds-averaged Navier--Stokes equations,
employing two different turbulence models---Menter's $k$--$\omega$ SST
\cite{Menter1994} and the Spalart--Allmaras (SA) one equation model
\cite{Spalart1992}. The SST model was chosen due to its prominence in the
literature for simulating separating flows, which we assumed to be present in
the current problem in the form of dynamic stall. The SA model was shown by
Ferreira et al. \cite{Ferreira2007} to match experimental particle image
velocimetry (PIV) results for a CFT in dynamic stall, though this was a somewhat
low Reynolds number case ($5 \times 10^4$).
